Probability Aggregation: Dynamic Hierarchical Modeling of Sparse Expert Beliefs

### 2013-09-04 paper.tex
Changed then name from STC --> SAC

%\usepackage[section]{placeins}
\usepackage{amsmath} 
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{fancyhdr}
\usepackage{moreverb}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{url}
\usepackage{multirow} 
\usepackage[boxed, section]{algorithm}
%\usepackage{algorithm}
\usepackage{algorithmic}
%\usepackage{cite}
\usepackage{multirow} 
\usepackage{rotating}
\usepackage{geometry}
\usepackage{fix-cm}
\usepackage{subfigure}
\usepackage{bbm}
\usepackage{color} 
\usepackage{natbib}
\usepackage{xcolor,xparse}
\usepackage{xparse}
%\usepackage{setspace}
\usepackage{booktabs}
%\usepackage{subfig}
%\usepackage{caption, subcaption}

@@

% Math Macros.  It would be better to use the AMS LaTeX package,
% including the Bbb fonts, but I'm showing how to get by with the most
% primitive version of LaTeX.  I follow the naming convention to begin
% user-defined macro and variable names with the prefix "my" to make it
% easier to distiguish user-defined macros from LaTeX commands.
%
\newcommand{\myN}{\hbox{N\hspace*{-.9em}I\hspace*{.4em}}}
\newcommand{\myZ}{\hbox{Z}^+}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\logit}{\text{logit}}
\NewDocumentCommand{\framecolorbox}{oommm}
 {% #1 = width (optional)
  % #2 = inner alignment (optional)
  % #3 = frame color
  % #4 = background color
  % #5 = text
  \IfValueTF{#1}
   {%
    \IfValueTF{#2}
     {\fcolorbox{#3}{#4}{\makebox[#1][#2]{#5}}}
     {\fcolorbox{#3}{#4}{\makebox[#1]{#5}}}%
   }
   {\fcolorbox{#3}{#4}{#5}}%
 }

@@

%\begin{keyword}[class=MSC]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

@@

%
%\title{Dynamic Hierarchical Model for Belief Streams}
%\author{by\\
%\\
%Ville A. Satop\"a\"a\\
%Shane T. Jensen and Lyle H. Ungar, Faculty Advisors\\
%\\
% \small Department of Statistics\\
% \small The Wharton School of the University of Pennsylvania\\
% \small Philadelphia, PA 19104- 6340, USA\\ [-0.25in]} \date{}
%\maketitle
%
%\pagestyle{myheadings}
%\markboth{Ville A. Satop\"a\"a}{Calibrated Crowd Belief}
%\thispagestyle{empty}

@@

\section{Model}
\label{model}
Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = 1, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$. Denote the logit-probabilities with 
\begin{eqnarray*}
Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R
\end{eqnarray*}
and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$. Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$. Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$. Let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the experts in question $k$; that is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector $\boldsymbol{e}_j$. The bias vector $\boldsymbol{b}$ does not include a subindex because it is considered shared among all the $K$ questions. To secure model identifiability, it is sufficient to share only one of the elements of $\boldsymbol{b}$ among the questions. This element defines a baseline under which it is possible to estimate the remaining $J-1$ bias terms separately within each of the questions. In this paper, however, the entire vector $\boldsymbol{b}$ is shared because some of the questions in our real-world data set involve very few experts with the highest level of self-reported expertise. Under this notation, the model for the $k$th question can be expressed as 
\begin{eqnarray}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} X_{t, k} + \boldsymbol{v}_{t, k} \label{observedpr} \\
X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k} \label{hiddenpr}\\
X_{0,k} &\sim& \mathcal{N}(\mu_0, \sigma^2_0) \nonumber
\end{eqnarray}
where Equation (\ref{observedpr}) denotes the observed process and Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$. The error terms are independent and identically distributed normal random variables with mean zero
\begin{eqnarray*}
\boldsymbol{v}_{t, k} | \sigma^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}(0, \tau^2_k),
\end{eqnarray*}
and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen \textit{a priori}. The hidden state $X_{t,k}$ represents the sharp and well-calibrated logit-probability for the $k$th event given the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. These probabilities are made estimable by assuming that the observed process is associated with the filtration $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ and hence can be used as a proxy for the amount of information available at any given time point. 
%Therefore it is not unreasonable to interpreted the hidden process as the crowd belief across time.

To give intuitive interpretations of the other model parameters, notice that the error term in the observed process has mean zero. Therefore the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief. An individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$. Unfortunately, due to the high sparsity of our data, estimating a bias term and analyzing the bias separately for each expert is not possible. See Section \ref{ExpertBias} for an analysis and discussion on the bias terms. Any other deviation from the calibrated crowd belief is considered \textit{random noise}. This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.

@@

The main challenge is to capture a well-calibrated estimate of the hidden process without sacrificing the interpretability of our model. This section introduces a two-step procedure, called \textit{Sample-Then-Calibrate} (STC), that achieves this goal in a flexible and efficient manner: the first step estimates the model parameters under a constraint (\textit{Sampling Step}), and the second step performs a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).

@@

\subsection{Calibration Step}
\label{calibration_step}
Since the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.  This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step. First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}. Since for any $\beta \in \R / \{0\}$,
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
 &=&   \boldsymbol{M}_k \left(\boldsymbol{b} (1) \beta \right) \left(X_{t, k}(1)/\beta\right)  + \boldsymbol{v}_{t, k}\\
&=&   \boldsymbol{M}_k \boldsymbol{b} (\beta) X_{t, k}(\beta) + \boldsymbol{v}_{t, k},
\end{eqnarray*}
the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
%Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$. Since the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$. A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}). Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions. Therefore,
\begin{eqnarray}
 \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
\end{eqnarray}
where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
\begin{eqnarray*}
S_{BRI}(Z, X) &=& -(Z - \logit^{-1}(X))^2
\end{eqnarray*}
or the logarithmic score (\citet{good1952rational})
\begin{eqnarray*}
S_{LOG}(Z, X) &=& Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)
\end{eqnarray*}
Since it is not clear which rule should be used for predicting geopolitical events, the \textit{Sample-Then-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}. Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
\begin{eqnarray}
 \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
 \hat{\boldsymbol{b}}_{t,k}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1)/ \hat{\beta}^2\nonumber\\
  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
\end{eqnarray}
Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint. Therefore their constrained and unconstrained versions are the same.

@@

\section{Synthetic Data Results}
\label{syntheticData}

The synthetic data is not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce well-calibrated hidden states. The latter is important for our interpretation of the hidden process as a sharp and well-calibrated version of the crowd belief. To ensure that the hidden process for question $k$ is well-calibrated, generate a path of the standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
\begin{eqnarray*}
Z_k &=& \mathbbm{1}\left( Z_{T_k, k}  > 0\right)\\
 X_{t,k} &=& \logit \left[ \Phi \left(\frac{Z_{t, k}}{\sqrt{T_k- t}} \right) \right]
\end{eqnarray*}
gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$.

We use this procedure to generate the hidden processes for $K$ questions with $T_k = 100$ for $k = 1, \dots, K$. Each question involves 50 experts allocated evenly among the five expertise groups. The expert logit-probabilities $Y_{i,t,k}$ are generated from Equation (\ref{observedpr}) by applying bias and noise to the hidden process $X_{t,k}$. Any extreme logit-probabilities are truncated to avoid overly influential observations; that is, any $p_{i,t,k} = \logit^{-1}(Y_{i,t,k})$ less than 0.01 (or greater than 0.99) is truncated to 0.01 (or to 0.99). This is akin to the real-world data setting, where extreme probabilities must be truncated (see Section \ref{realData}). Therefore, to understand the effect that truncation has on the real-world data results, truncation should be first analyzed and understood under the controlled environment of synthetic data.

Throughout the evaluation, data is generated under different choices of parameter values. More specifically, the evaluation procedure iterates over a three-dimensional grid of parameter values
\begin{eqnarray*}
\beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
\sigma^2 &\in& \{0.5, 1.0, 1.5, 2.0, 2.5\}\\
K &\in& \{20, 35, 50, 65, 80\},
\end{eqnarray*}
where $\beta$ generates the bias vector  $\boldsymbol{b} = [0.50, 0.75, \dots, 1.50]^T \beta$. These values were considered to be realistic based on our experience with the real-world probability forecasting data. Each triple of the parameter values,  i.e. a grid point, is used  40 times to generate a synthetic dataset. Each dataset is then used to assess how accurately \textit{Sample-Then-Calibrate} is able to estimate the hidden process and bias vector. The accuracy of the estimated hidden process is measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$, and the accuracy of the estimtated bias vector is quantified by the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.

@@

\caption[Optional caption for list of figures]{Comparing \textit{Sample-Then-Calibrate} that optimizes over the logarithmic score (STC-Log) and \textit{Sample-Then-Calibrate} that optimizes over the Brier score (STC-Bri) under synthetic data. The top row presents the accuracy to capture the hidden process. The bottom row presents the accuracy to capture the bias vector. }
\label{Synthetic}
\end{figure}

Figure \ref{Synthetic} summarizes the results in a series of plots. The first row shows the marginal effects of the three grid variables on the accuracy to which the \textit{Sample-Then-Calibrate} under the logarithmic loss (STC-Log) and the Brier score (STC-Bri) are able to estimate the hidden process.  The second row is similar in nature but measures the accuracy of estimating the bias vector instead.  The plots were computed by choosing one grid variable (e.g. $\beta$) at a time and averaging over the remaining two variables  (e.g. $K$, and $\sigma^2$).  Therefore each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.

Based on the first row of plots, STC-Log and STC-Bri estimate the hidden process to a similar accuracy. Even though some relative differences are visible, the loss values are very small in absolute terms, hence suggesting good overall accuracy to capture the hidden process. Estimating the hidden process, however, becomes more difficult as either $\sigma^2$ or $\beta$ gets larger. The latter seems more unintuitive and can be understood better by observing that larger values of $\beta$ lead to more extreme expert logit-probabilities and hence to more frequent truncation. This causes the estimation accuracy to degrade because  a high level of truncation results to a drastic deviation from the model assumptions. Increasing the number of questions improves the estimation accuracy but only very little.

Based on the second row of plots, STC-Log and STC-Bri  capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very large. This behavior can be again considered as being caused by the truncation of the extreme expert logit-probabilities. Surprisingly, even though STC-Log is more accurate in estimating the bias vector, it is unable to outperform STC-Bri in estimating the hidden process.

\section{Geopolitical Data Results}
\label{realData}
\noindent
This section discusses results related to the real-world data described in Section \ref{data}. The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro} and also to evaluate the \textit{Sample-Then-Calibrate}-procedure in terms of predictive power and calibration. Before presenting the results, however, we discuss two practical matters that must be taken into account when aggregating real-world  probability forecasts.

>>>>

%\usepackage[section]{placeins}
\usepackage{amsmath} 
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{fancyhdr}
\usepackage{moreverb}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{url}
\usepackage{multirow} 
\usepackage[boxed, section]{algorithm}
%\usepackage{algorithm}
\usepackage{algorithmic}
%\usepackage{cite}
\usepackage{multirow} 
\usepackage{rotating}
%\usepackage[margin=1.2in]{geometry}
\usepackage{geometry}
\usepackage{fix-cm}
\usepackage{subfigure}
\usepackage{bbm}
\usepackage{color} 
\usepackage{natbib}
\usepackage{xcolor,xparse}
\usepackage{xparse}
%\usepackage{setspace}
\usepackage{booktabs}
%\usepackage{subfig}
%\usepackage{caption, subcaption}

@@

% Math Macros.  It would be better to use the AMS LaTeX package,
% including the Bbb fonts, but I'm showing how to get by with the most
% primitive version of LaTeX.  I follow the naming convention to begin
% user-defined macro and variable names with the prefix "my" to make it
% easier to distiguish user-defined macros from LaTeX commands.
%
\newtheorem{mydef}{Definition}
\newcommand{\myN}{\hbox{N\hspace*{-.9em}I\hspace*{.4em}}}
\newcommand{\myZ}{\hbox{Z}^+}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\logit}{\text{logit}}
\NewDocumentCommand{\framecolorbox}{oommm}
 {% #1 = width (optional)
  % #2 = inner alignment (optional)
  % #3 = frame color
  % #4 = background color
  % #5 = text
  \IfValueTF{#1}
   {%
    \IfValueTF{#2}
     {\fcolorbox{#3}{#4}{\makebox[#1][#2]{#5}}}
     {\fcolorbox{#3}{#4}{\makebox[#1]{#5}}}%
   }
   {\fcolorbox{#3}{#4}{#5}}%
 }

@@

@@

@@

\section{Model}
\label{model}
Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = 1, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$. Denote the logit-probabilities with 
\begin{eqnarray*}
Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R
\end{eqnarray*}
and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$. Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$. Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$. Let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the experts in question $k$; that is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector $\boldsymbol{e}_j$. The bias vector $\boldsymbol{b}$ does not include a subindex because it is considered shared among all the $K$ questions. To secure model identifiability, it is sufficient to share only one of the elements of $\boldsymbol{b}$ among the questions. This element defines a baseline under which it is possible to estimate the remaining $J-1$ bias terms separately within each of the questions. In this paper, however, the entire vector $\boldsymbol{b}$ is shared because some of the questions in our real-world data set involve very few experts with the highest level of self-reported expertise. Under this notation, the model for the $k$th question can be expressed as 
\begin{eqnarray}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} X_{t, k} + \boldsymbol{v}_{t, k} \label{observedpr} \\
X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k} \label{hiddenpr}\\
X_{0,k} &\sim& \mathcal{N}(\mu_0, \sigma^2_0) \nonumber
\end{eqnarray}
where Equation (\ref{observedpr}) denotes the observed process and Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$. The error terms are independent and identically distributed normal random variables with mean zero
\begin{eqnarray*}
\boldsymbol{v}_{t, k} | \sigma^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}(0, \tau^2_k),
\end{eqnarray*}
and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen \textit{a priori}.

The hidden state $X_{t,k}$ represents the logit-probability for the $k$th event given all the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. It is reasonable to assume that an expert, who has access to all available information and reports $\logit^{-1}(X_{t,k})$ as the probability forecast, maximizes sharpness subject to calibration, where calibration and sharpness are understood as follows:
\begin{mydef}
A probability forecast $p$ for the $k$th question is calibrated if $\P(Z_k = 1 | p) = \E[Z_k | p] = p$ almost surely (see, e.g., \citet{murphy1987general}).
\end{mydef}
\begin{mydef}
A probability forecast $p$ is sharper than a probability forecast $q$ if $\E[(p-p_0)^2] > \E[(q-p_0)^2]$, where $p_0$ is the baseline probability for the $k$th question (see, e.g., \citet{ranjan2009combining}). 
\end{mydef}
\noindent
Even though it is unlikely that any single expert has all the available information, a large and diverse group of experts may share a considerable portion of this information. The so called \textit{wisdom of the crowd} is therefore an attractive proxy for the amount of information available at any given time point.

If a model is to efficiently use this proxy for estimating the hidden process $X_{t,k}$, it must be able to discriminate between noise and signal as experts may believe in false information, hide their true beliefs on an ethical basis, or be biased for many other reasons. Since in our model the experts are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief, an individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$.  Due to the high sparsity of our data, estimating a bias term and analyzing the bias separately for each expert is not possible. See Section \ref{ExpertBias} for an analysis and discussion on the bias terms. Any other deviation from the calibrated crowd belief is considered \textit{random noise}. This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.

@@

The main challenge is to capture a well-calibrated estimate of the hidden process without sacrificing the interpretability of our model. This section introduces a two-step procedure, called \textit{Sample-And-Calibrate} (SAC), that achieves this goal in a flexible and efficient manner: the first step estimates the model parameters under a constraint (\textit{Sampling Step}), and the second step performs a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).

@@

\subsection{Calibration Step}
\label{calibration_step}
Since the model parameters can be estimated by fixing $b_3$ to any constant, the next step is to search for the constant that gives an optimally sharp and calibrated estimate of the hidden process.  This section introduces an efficient procedure that finds the optimal constraint without requiring any additional runs of the sampling step. First, assume that parameter estimates  $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ have already been obtained via the constrained sampling step described in Section \ref{sampling_step}. Since for any $\beta \in \R / \{0\}$,
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
 &=&   \boldsymbol{M}_k \left(\boldsymbol{b} (1) \beta \right) \left(X_{t, k}(1)/\beta\right)  + \boldsymbol{v}_{t, k}\\
&=&   \boldsymbol{M}_k \boldsymbol{b} (\beta) X_{t, k}(\beta) + \boldsymbol{v}_{t, k},
\end{eqnarray*}
the parameter values under $b_{3} = \beta$ can be obtained from $\boldsymbol{b} (\beta) = \boldsymbol{b} (1) \beta $ and $X_{t, k}(\beta) = X_{t, k}(1)/\beta$. 
%Furthermore, an estimate of $X_{t, k}(\beta)$ is given by $\hat{X}_{t, k}(1)/\beta$. 
This means that $X_{t,k} = X_{t,k}(1)/\beta$ when $\beta$ is equal to the true value of $b_3$. Since the hidden process $X_{t, k}$ is assumed to be sharp and well-calibrated, $b_3$ can be estimated with the value of $\beta$ that simultaneously maximizes the sharpness and calibration of $\hat{X}_{t, k}(1) / \beta$. A natural criterion for this maximization is given by the class of \textit{proper scoring rules} that combine sharpness and calibration (\citet{gneiting2008rejoinder, buja2005loss}). Due to the possibility of \textit{complete separation} in any one question (see, e.g., \citet{gelman2008weakly}), the maximization must be performed over multiple questions. Therefore,
\begin{eqnarray}
 \hat{\beta}&=&  \argmax_{\beta \in \R / \{0\}} \sum_{k=1}^K \sum_{t=1}^{T_k}  S\left(Z_k, \hat{X}_{k,t}(1) / \beta \right) \label{OSE}
\end{eqnarray}
where $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. The function $S$ is a strictly proper scoring rule such as the negative Brier score (\citet{Brier})
\begin{eqnarray*}
S_{BRI}(Z, X) &=& -(Z - \logit^{-1}(X))^2
\end{eqnarray*}
or the logarithmic score (\citet{good1952rational})
\begin{eqnarray*}
S_{LOG}(Z, X) &=& Z \log\left(\logit^{-1}(X)\right) + (1-Z) \log\left(1-\logit^{-1}(X)\right)
\end{eqnarray*}
Since it is not clear which rule should be used for predicting geopolitical events, the \textit{Sample-And-Calibrate} procedure is evaluated  separately under both rules in Sections \ref{syntheticData} and \ref{realData}. Once $\hat{\beta}$ has been computed, estimates of the unconstrained model parameters are given by
\begin{eqnarray}
 \hat{X}_{t,k}&=& \hat{X}_{k,t}(1) / \hat{\beta} \nonumber\\
 \hat{\boldsymbol{b}}_{t,k}&=& \hat{\boldsymbol{b}}(1) \hat{\beta} \nonumber\\
   \hat{\tau}_{k}^2&=& \hat{\tau}_{k}^2(1)/ \hat{\beta}^2\nonumber\\
  \hat{\sigma}_{k}^2&=& \hat{\sigma}_{k}^2(1)\nonumber\\
  \hat{\gamma}_{k}&=& \hat{\gamma}_{k}(1)\nonumber
\end{eqnarray}
Notice that estimates of $\sigma^2_k$ and $\gamma_k$ are not affected by the constraint. Therefore their constrained and unconstrained versions are the same.

@@

\section{Synthetic Data Results}
\label{syntheticData}
The synthetic data is not generated directly from the model for two reasons: (i) showing good performance on a dataset directly generated from the model assumptions is hardly any news, and (ii) generating data from the dynamic model description does not produce well-calibrated hidden states. The latter is important for our interpretation of the hidden process as a sharp and well-calibrated version of the crowd belief. To ensure that the hidden process for question $k$ is well-calibrated, generate a path of the standard Brownian motion until time $T_k$. If $Z_{t,k}$ denotes the value of the path at time $t$, then
\begin{eqnarray*}
Z_k &=& \mathbbm{1}\left( Z_{T_k, k}  > 0\right)\\
 X_{t,k} &=& \logit \left[ \Phi \left(\frac{Z_{t, k}}{\sqrt{T_k- t}} \right) \right]
\end{eqnarray*}
gives a sequence of $T_k$ calibrated logit-probabilities for the event $Z_k = 1$. Repeating this procedure for $K$ times, we generate the hidden process for $K$ questions, each with a time horizon of $T_k = 101$. The questions involve 25 experts allocated evenly among five expertise groups. Since the event resolves at time $t = 101$, the expert logit-probabilities $Y_{i,t,k}$ are only given for times $t = 1$ to $100$. These are generated from Equation (\ref{observedpr}) by applying bias and noise to the hidden process $X_{t,k}$. The simulation then iterates over a three-dimensional grid of parameter values
\begin{eqnarray*}
\sigma^2 &\in& \{1/2, 1, 3/2, 2, 5/2\}\\
\beta &\in& \{1/2, 3/4, 1, 4/3, 2/1\}\\
K &\in& \{20, 40, 60, 80, 100\},
\end{eqnarray*}
where $\beta$ gives the value of the bias vector by letting $\boldsymbol{b} = [1/2, 3/4, 1, 4/3, 2/1]^T \beta$. Each grid point is used  40 times to generate a synthetic dataset. This dataset is then used to assess how accurately \textit{Sample-And-Calibrate} estimates the hidden process and bias vector. The accuracy of the estimated hidden process is measured with the average quadratic loss in the probability space, $\sum_{k=1}^K \sum_{t=1}^{T_k} ( \logit^{-1}(\hat{X}_{t,k}) - \logit^{-1}(X_{t,k}))^2 / \sum_{k=1}^K T_k$, and the accuracy of the estimated bias vector is measured with the average quadratic loss, $|| \boldsymbol{b} - \hat{\boldsymbol{b}}||^2/5$.

@@

\caption[Optional caption for list of figures]{Comparing \textit{Sample-And-Calibrate} that optimizes over the logarithmic score ($\text{SAC}_{\text{LOG}}$) and \textit{Sample-And-Calibrate} that optimizes over the Brier score (SAC-Bri) under synthetic data. The top row presents the accuracy to capture the hidden process. The bottom row presents the accuracy to capture the bias vector. }
\label{Synthetic}
\end{figure}

Figure \ref{Synthetic} summarizes the results in a series of plots. The first row shows the marginal effects of the three grid variables on the accuracy to which the \textit{Sample-And-Calibrate} under the logarithmic loss ($\text{SAC}_{\text{LOG}}$) and the Brier score (SAC-Bri) are able to estimate the hidden process.  The second row is similar in nature but measures the accuracy of estimating the bias vector instead.  Since the plots were computed by choosing one grid variable (e.g. $\beta$) at a time and averaging over the remaining two variables  (e.g. $K$, and $\sigma^2$), each value in the plots represents an average of a $40 \times 5^2 = 1,000$ values.

Based on the first row of plots, $\text{SAC}_{\text{LOG}}$ and SAC-Bri estimate the hidden process to a similar accuracy. Even though some relative differences are visible, the loss values are very small in absolute terms, hence suggesting good overall accuracy to capture the hidden process. Estimating the hidden process, however, becomes more difficult as either $\sigma^2$ or $\beta$ gets larger. The latter seems more unintuitive and can be understood better by observing that larger values of $\beta$ lead to more extreme expert logit-probabilities and hence to more frequent truncation. This causes the estimation accuracy to degrade because  a high level of truncation results to a drastic deviation from the model assumptions. Increasing the number of questions improves the estimation accuracy but only very little.

Based on the second row of plots, $\text{SAC}_{\text{LOG}}$ and SAC-Bri  capture the bias vector rather precisely as long as the true values of $\beta$ and $\sigma^2$ are not very large. This behavior can be again considered as being caused by the truncation of the extreme expert logit-probabilities. Surprisingly, even though $\text{SAC}_{\text{LOG}}$ is more accurate in estimating the bias vector, it is unable to outperform SAC-Bri in estimating the hidden process.

\section{Geopolitical Data Results}
\label{realData}
\noindent
This section presents results related to the real-world data described in Section \ref{data}. The goal is to provide application specific insight by discussing the specific research objectives itemized in Section \ref{intro} and also to evaluate the \textit{Sample-And-Calibrate}-procedure in terms of predictive power and calibration. Before presenting the results, however, we discuss two practical matters that must be taken into account when aggregating real-world  probability forecasts.

\subsection{Incoherent and Imbalanced Data}
\label{practicalMatters}
The first matter regards human experts making probability forecasts of 0.0 or 1.0 even if they are not completely sure of the outcome of the event. For instance, all the 166 questions in our dataset contained both a zero and a one. Transforming such forecasts into the logit-space yields infinities that are highly influential and cause problems in model estimation. To avoid this, \citet{Ariely00theeffects} suggest changing $p = 0.00$ and $1.00$ to $p = 0.02$ and $0.98$, respectively. This is similar to \textit{winsorising} that sets the extreme probabilities to a specified percentile of the data (see, e.g., \citet{hastings1947low} for more details on winsorising).  \citet{Geo}, on the other hand, consider only probabilities that fall within a constrained interval, say $[0.001, 0.999]$, and discard the rest. Since this implies ignoring a portion of the data, we adopt the first approach by truncating $p = 0.00$ and $1.00$ to $p = 0.01$ and $0.99$, respectively.  Our results remain insensitive to the exact choice of truncation as long as this is done in a reasonable manner to keep the extreme probabilities from becoming highly influential in the logit-space.
